---
title: "Network analysis with Network D3"
author: "Paul Bradshaw"
date: "11 February 2017"
output: html_document
---

# Creating network graphs with [NetworkD3](https://christophergandrud.github.io/networkD3/). You can [find examples on the GitHub](https://github.com/christophergandrud/networkD3/blob/master/inst/examples/examples.R)

```{r}
#Install the networkD3 package, and activate the library
install.packages('networkD3')
library(networkD3)
```

```{r}
#Import the dataset
chars <- read.csv('starwars/swpeople.csv')
#Put each column into its own list
films <- chars$film.name
names <- chars$X..name
```

Now we need to create the nodes to generate relationships - and then generate a network from those two

```{r}
src <- names
target <- films
networkData <- data.frame(src, target)
```

Of course we could have just done the same by grabbing the column directly like so:

`networkData <- data.frame(chars$film.name, chars$X..name)`

Then draw the network diagram:

```{r}
simpleNetwork(networkData)
```

And save it as a HTML file:

```{r}
library(magrittr)

simpleNetwork(networkData) %>%
saveNetwork(file = 'swpeople.html')
```

This is still a very simple network diagram - there is no indication of the strength of relationships or which nodes appear most. 

```{r}
# Create graph with legend and varying radius
forceNetwork(Links = MisLinks, Nodes = MisNodes, Source = "source",
             Target = "target", Value = "value", NodeID = "name",
             Nodesize = 'size', radiusCalculation = "d.nodesize",
             Group = "group", opacity = 1, legend = T, bounded = F) %>%
saveNetwork(file = 'forcenet.html')
```

Let's reverse-engineer this a little. Each data frame has 3 columns. `MisNodes` is obviously the nodes, and `MisLinks` is presumably the 'edges' - the links between nodes.

Here's a hypothesis: the links dataset only contains numbers. So it can't be drawing links based on the names of each item. 

The nodes dataset has two numerical columns: *group* and *size*. Group is shown in the colour coding. And size appears to determine the size of each node in the diagram. So perhaps the *index* of each row - the row number - is what the links are using.

The pink items in group 7 are a little isolated. We can change one of the links like so to test if the link is affected. Before this value was 46, but we change it to 47 so it links to itself:

```{r}
#Show the values in row 109
MisLinks[109,]
#Show the value in the second cell in that row
MisLinks[109,2]
#Change the value
MisLinks[109,2] <- 47
```

And then re-draw the network. 

```{r}
# Create graph with legend and varying radius
forceNetwork(Links = MisLinks, Nodes = MisNodes, Source = "source",
             Target = "target", Value = "value", NodeID = "name",
             Nodesize = 'size', radiusCalculation = "d.nodesize",
             Group = "group", opacity = 1, legend = T, bounded = F) %>%
saveNetwork(file = 'forcenet3.html')
```

Note that one node is now *on its own* - it's the one in *row* 47 in the node data frame, *Jondrette*

What about the *value* in the MisLinks data frame? Let's change one from 1 to 10 to see:

```{r}
MisLinks[109,3] <- 10
```

Look closely and you'll see one of the links is a lot thicker. In fact, you need to change it to at least 100 to really see:

```{r}
MisLinks[109,3] <- 100
```

So now we can conclude:

* In MisNodes, *label* is used to label each node
* *group* is used to classify each node, and colour code them accordingly if needed
* *size* is used to control the size of each node, if needed
* The *index* of each element will be used to connect them
* In MisLinks, *source* begins drawing a line from one element: it uses the node's row number (index)
* *target* ends the line, normally at the index of a different element. If it is the same, no line is drawn
* *size* determines the thickness of the line, at least in the code that we then use.

Create a subset to test this a bit further

```{r}
MisLinks3 <- MisLinks[c(1:10),]
```

Most of these link to 0 - and this turns out to be the entry in row 1. So we have a 0-based index.

