---
title: "Network analysis with Network D3"
author: "Paul Bradshaw"
date: "11 February 2017"
output: html_document
---

# Creating network graphs with [NetworkD3](https://christophergandrud.github.io/networkD3/). You can [find examples on the GitHub](https://github.com/christophergandrud/networkD3/blob/master/inst/examples/examples.R)

```{r}
#Install the networkD3 package, and activate the library
install.packages('networkD3')
library(networkD3)
```

```{r}
#Import the dataset
chars <- read.csv('starwars/swpeople.csv')
#Put each column into its own list
films <- chars$film.name
names <- chars$X..name
```

Now we need to create the nodes to generate relationships - and then generate a network from those two

```{r}
src <- names
target <- films
networkData <- data.frame(src, target)
```

Of course we could have just done the same by grabbing the column directly like so:

`networkData <- data.frame(chars$film.name, chars$X..name)`

Then draw the network diagram:

```{r}
simpleNetwork(networkData)
```

And save it as a HTML file:

```{r}
library(magrittr)

simpleNetwork(networkData) %>%
saveNetwork(file = 'swpeople.html')
```

This is still a very simple network diagram - there is no indication of the strength of relationships or which nodes appear most. 
